---
title: "RK_607_Project01"
author: "Raj Kumar"
date: "February 22, 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## PROJECT 01: Chess Player Rating Ingestion Tool


### INDEX (Step by Step)
#####    STEP 1. Load Libraries
#####    STEP 2. Load the file
#####    STEP 3. Data Extraction:Regex the file data 
#####    STEP 4. Data Munging: Create measurement of opponents score
#####    STEP 5. Data Munging: Create a final dataframe as suggested in the problem
#####    STEP 6. Create interactive Data Table from the dataframe
#####    STEP 7. Dump the data into CSV file
#####    STEP 8. Use the SQL script to create new chess table and load newly created csv file into the chess table
#####    STEP 9. Query the chess table from this RMD and check the results
#####    STEP 10. success!!


### IMPORTANT PROJECT ASSUMPTIONS 
#####    Assumption 1. Pre Tournament rating was considered as player's rating for final table
#####    Assumption 2. For Opponent avg. score calculation: 
#####                   a. All opponents numbers were matched with their numbers in the file
#####                   b. The codes like U and H where no value for opponent was given in the file were ignored from the calculation in order to make sure data is valid
#####    Assumption 3. Rating (example 1220P13) was treated as rating of 1220 as meaning of P13 was not clear. 


## STEP 1 : Load your libraries
```{r}
# Load the libraries
library(stringr)    #For Regex
library(RCurl)      #For File Operations
library(dplyr)      #For Manipulating the data frames
library(DT)         #For Data table package
library(RMySQL)     #For SQL Operations
library(ggplot2)    #For Visualizations

```

## STEP 2 : Load the File
```{r}
# Good Practise: Basic house keeping: cleanup the env before you start new work
rm(list=ls())

# Good Practise: Set up the Working Directory when working with a file system
setwd("C:\\CUNY\\607Data\\Assignments\\project01")

# Read the File
raw.file.data <- readLines("tournamentinfo.txt")

# Make sure file loaded correct
head(raw.file.data)
```


## STEP 3 : Data Extraction:Regex the file data 

#### Regex to find Player Names
```{r}

##
# Now the cleanup fun begins
##

##  
#   I created my regex and tested these using website 
#   http://rubular.com/
##


##  
#   Regex name of the players
##
namenum        <- unlist(str_extract_all(raw.file.data, 
                                      "\\d{1,2}[ ]\\|[ ]([A-Z]+[ ]){2,3}"))
name        <- unlist(str_extract_all(namenum, "([A-Z]+[ ]){2,3}"))
name        <- trimws(name, which = c("both"))
name

```


#### Regex to find Player Numbers
```{r}
##  
#   Regex number of the players
##
pnum   <- as.numeric(unlist(str_extract_all(namenum, "\\d{1,2}")))
pnum

```


#### Regex to find Player Pre Tournament Rating
```{r}
##  
#   Regex players rating
##
rating      <- unlist(str_extract_all(raw.file.data, "R:[ ]+(\\d{3,4})"))
rating      <- as.numeric(unlist(str_extract_all(rating, "\\d{3,4}")))
rating


```



#### Regex to find Player State
```{r}
##  
#   Regex players state
##
state     <- unlist(str_extract_all(raw.file.data, "\\b([A-Z]{2})[ ]\\|"))
state     <- unlist(str_extract_all(state, "[A-Z]{2}"))
state



```


#### Regex to find Player Points
```{r}
##  
#   Regex the points
##
points      <- as.numeric(unlist(str_extract_all(raw.file.data, "\\d\\.\\d")))
points

```


#### Create a Dataframe with the data
```{r}
pointsdf <- data.frame(pnum=pnum, name=name, state=state, points=points, rating=rating)
head(pointsdf)


```



## STEP 4: Create measurement of opponents scores

#### Cleanup the file further to find list of opponents for each player
```{r}
## Cleanup lines to derive the average of opponents

# Remove the lines that just contain ------- 
raw.data <- str_extract_all(raw.file.data, "^(?!.*--.*).+$")

# Create a list of opponents each player played
opponentslist <-  str_extract_all(raw.data,"[WLDHU][ ]+\\d{1,2}")

# remove any empty rows from the list
opponentslist <- opponentslist[lapply(opponentslist,length)>0]

#check the list before processing futher
opponentslist


```


#### For loop to compute opponent averages for all the players
```{r}
# Compute the average value for all the opponents

avg_list <- c()
for (i in 1:64){
        sum <- 0
        for (j in 1:length(opponentslist[[i]])){
            opponent <- as.numeric(str_extract_all(
                opponentslist[[i]][j], "\\d{1,2}"))            
                sum <- sum + as.numeric(as.character(
                    pointsdf$rating[which(pointsdf$pnum==opponent)][1]))
        }
        #print(sum)
        avg_list[i] <- round(sum/length(opponentslist[[i]]))
}
#all opponent averages
avg_list

```


#### Add the opponent averages to the existing data frame
```{r}
# Add the opponext average to the existing data frame
pointsdf["opponentAvg"] <- avg_list

```


## STEP 5: Create a final dataframe as suggested in the problem
```{r}
##
# Cleanup and format the dataframe 
##

pointsdf <- subset(pointsdf, select = c(name, state, points, rating, opponentAvg))

format(pointsdf$points, digits=1, nsmall=1)

head(pointsdf)

```


## STEP 6: Create a Beautiful Data Table for the data frame 
```{r}
datatable(pointsdf)

```

## STEP 7: Write the dataframe to CSV file
```{r}
## Write the file in CSV format to the disk
write.table(pointsdf, "chessdata.csv", sep=",", row.names=FALSE, col.names=FALSE)
```

## STEP 8: SQL SCript to create chess table
```{r}
# Create a database connection
mydb = dbConnect(MySQL(), user='root', password='mysql57', host='localhost', dbname="tb")

```

#### Exhibits for my SQL Scripts to create tables
```{ }
/*
 The below are just exhibits from the SQL Script which I ran using mysql databases
 This Script create the table in the database
 The actual script is attached with this project RMD code 
 
 A. Create Table
 B. Insert the rows using the CSV created by the above write.csv command
 C. Check the results
*/

/*
DROP TABLE IF EXISTS chess;

# Create new table
CREATE TABLE chess 			
(
  name varchar(100) NOT NULL,  
  state varchar(10) NOT NULL,  
  points float4 NOT NULL,  
  rating int NOT NULL,  
  opponentavg int NOT NULL	 
);

LOAD DATA INFILE 'C:\\ProgramData\\MySQL\\MySQL Server 5.7\\Uploads\\chessdata.csv'  
INTO TABLE chess 
FIELDS TERMINATED BY ','  
lines terminated BY "\n";

*/
```



## STEP 9: Database connection to check SQL operations
```{r}
##
# Create a Database connection
##

mysqloutput <- fetch(dbSendQuery(mydb, "select * from chess;"))

head(mysqloutput)
```

## STEP 10: Success!!


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
