---
title: "RK_607_Assignment02_DB"
author: "Raj Kumar"
date: "February 9, 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## Assignment - SQL and R
### Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.
###Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.
###This is by design a very open ended assignment. A variety of reasonable approaches are acceptable. You can (and should) blank out your SQL password if your solution requires it; otherwise, full credit requires that your code is "reproducible," with the assumption that I have the same database server and R software.


### I choose a combination of English and Hindi movies that our family had watched recently 
### Movie Names: Bahubali2, Dunkrik, Judwaa2, Jumanji2, Mubarakan and Padmavaat

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup the Table 
###I created a table 'movies' with three columns
###    1. mnane: movie name as varchar(100) not null
###    2. fname: friend name as varchar(100) not null
###    3. rating: ratings for the movie as an int
### Please see the SQL script (movie.sql) for more details

    
1.  Load the Libraries  

### Load the required libraries
```{r}
## Load the Libraries 
library(RMySQL)
library(odbc)
library(ggplot2)

```

    
2.  Connect to the My SQL Database  

### Connect to mySQL Database
```{r}

## connect using dbConnect
mydb = dbConnect(MySQL(), user='root', password='mysql57', host='localhost', dbname="tb")

```


3. Query the Data base 

### Query the database
```{r}
moviedf = fetch(dbSendQuery(mydb, "select * from movies;"))

```

4. Review the results from the memory

### Review the results 
```{r}
dim(moviedf)
moviedf

```

5.  Analysis of the data for each movie

### Data Analysis for each movie 

```{r}

#Print the rating for each movie and find the means rating 


moviedf$rating[which(moviedf$mname=="Bahubali2")]
m1 = mean(moviedf$rating[which(moviedf$mname=="Bahubali2")])
m1

moviedf$rating[which(moviedf$mname=="Dunkrik")]
m2 = mean(moviedf$rating[which(moviedf$mname=="Dunkrik")])
m2 

moviedf$rating[which(moviedf$mname=="Judwaa2")]
m3 = mean(moviedf$rating[which(moviedf$mname=="Judwaa2")])
m3

moviedf$rating[which(moviedf$mname=="Jumanji2")]
m4 = mean(moviedf$rating[which(moviedf$mname=="Jumanji2")])
m4

moviedf$rating[which(moviedf$mname=="Mubarakan")]
m5 = mean(moviedf$rating[which(moviedf$mname=="Mubarakan")], na.rm=TRUE)
m5

moviedf$rating[which(moviedf$mname=="Padmavaat")]
m6 = mean(moviedf$rating[which(moviedf$mname=="Padmavaat")], na.rm = TRUE)
m6





```
### Lets find the Best and Worst movie... Drum rolls!!!
```{r}
allrating <- c(m1, m2, m3, m4, m5, m6) 
names(allrating) <- names(table(moviedf$mname))

# And the highest rated movie was .... drum roll!!!
allrating[which(allrating==max(allrating))]


# And the lowest rated movie was .... booooooo!!!
allrating[which(allrating==min(allrating))]

```


6. Analysis using GGPLOT2 
### let draw a GGPlot for the ratings
```{r}
# Draw a GGPlot to show how each movie was rated 
# I am using Box Plot as it can show the range of reviews and provide a mean for these ratings
ggplot(moviedf, aes(mname, rating))+
    geom_boxplot(na.rm = TRUE, color = 'blue', outlier.color = 'green')+
    theme_bw()+
    labs(title="Movie Review", x="Movie Name", y="Star Ratings")+
    theme(plot.title = element_text(hjust = 0.5))


```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
