---
title: "RK_607_Assignment03_REGEX"
author: "Raj Kumar"
date: "February 9, 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## Assignment 03 - REGEX
### Here is the referenced code for the introductory example in #3:
### raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert" 


## Character Data Exploration with REGEX

### STEP 1 : Always load your libraries
```{r}
library(stringr)
```

### STEP 1 : Data Setup 
```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone

data.frame(name=name, phone=phone)

```


1. Copy the introductory examples. The Vector names stores the extracted names

###(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.
```{r}
name

tempname <- str_replace_all(name, "(.+)(, .+)$", "\\2 \\1") 
tempname <- str_replace_all(tempname, ", ", "") 
tempname <- str_replace_all(tempname, "[A-Z][a-z]([a-z]?)\\.", "") 
tempname

# Create lists with first name and last name
fname <- str_extract(tempname, "[[:alpha:]+]{2,}")
lname <- str_extract(tempname, "[[:space:]][[:alpha:]+]{2,}")
fname
lname

namedf <- data.frame(fname=fname, lname=lname)
namedf
```


###(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r}
hasTitle <- str_detect(name, "Rev.|Dr.")
hasTitle
```


###(c) Construct a logical vector indicating whether a character has a second name.
```{r}
hasSecondName <- str_detect(name, "[A-Z]\\.")
hasSecondName
```


2. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

###(a) Any string with $ sign
```{r }
pattern <-  "[0-9]+\\$"
list=c("555$","ab12$USD",".01$","23213.121")
str_detect(list,pattern)
```

###(b) Any string with a-z only
```{r }
pattern <-  "\\b[a-z]{1,4}\\b"
list=c("a","bb212", "xyzbc","xz","aaa","12abcdefgh12a", "1234")
str_detect(list,pattern)

```

###(c) match string with .txt ending
```{r }
pattern <-  ".*?\\.txt$"
list=c("a.txt","bb212", "xyzbc.txt","xz","aaa.txt","12abcdefgh12a", "1234.txt")
str_detect(list,pattern)

```


###(d) Match number in data format
```{r }
pattern <-  "\\d{2}/\\d{2}/\\d{4}"
list=c("10/10/2015 Check Date", "20202015", "someone2018/02/07", "99999999", "99/99/9999")
str_detect(list,pattern)

```

###(e) Tag Format
```{r}
pattern <-  "<(.+?)>.+?</\\1>"
list=c("<b>name</b>", "<a href=testurl>link</a>", "324324 sdsadd 213213","<xml><head></head></xml>", "99999999")
str_detect(list,pattern)
```


3. The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com

```{r}
##
# The message clearly is written in capital, so we can just extract everything in capital
##
message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

unlist(str_extract_all(message, "[[:upper:].!]"))

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
