---
title: "RK_607_Week09_Working_with_Web_APIs"
author: "Raj Kumar"
date: "March 25th, 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

<div class="jumbotron">
  <h2 class="display-4">Web API</h2>
  <p class="lead">
The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs
You'll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.
</p>
  <hr class="my-4">
</div>


### INDEX (Step by Step)
#####    STEP 1. Setup the API key
#####    STEP 2. Load the libraries
#####    STEP 3. Read the data
#####    STEP 4. Convert the data into usable dataframe 


## STEP 0 : Cleanup the Environment
```{r echo=FALSE, message=FALSE, results=FALSE}
# Good Practise: Basic house keeping: cleanup the env before you start new work
rm(list=ls())

# Garbage collector to free the memory
gc()

# Good Practise: Set the graphics drivers 
dev.off()

# Good Practise: Set up the Working Directory when working with a file system
setwd("C:\\CUNY\\607Data\\Assignments")

```

## STEP 1 : Set up the API Keys
New York times API can be registered to at http://developer.nytimes.com/signup




## STEP 2 : Load your libraries
```{r}
# Load the libraries
library("httr")
library("rjson")
library("DT")
```


##    STEP 3. Read the data
```{r}
api.key <- "0bed27e078464cd595721a8cd5243dd5"
url <- "http://api.nytimes.com/svc/topstories/v1/home.json?api-key="

result <- GET(paste(url, api.key, sep=""))

```



##    STEP 4. Convert the data into usable dataframe 
```{r}
url.content <- content(result)

content.list <- fromJSON(url.content)
content.df <- as.data.frame(do.call(rbind, content.list$results))
colnames(content.df)

datatable(content.df)

```

